---
title: "MOTUS telemetry array to access habitat selection from migratory shorebirds within and across Hunter and Port Stephen estuaries, NSW"
date:  last-modified
date-format: "MMMM D, YYYY"
css: styles.css
format:
  html:
    code-line-numbers: true
project:
  type: website
  output-dir: 22_qmd
---

<p style="font-size: 0.7em; color: gray;">Created on July 28, 2025</p>

## [*Maxime Marini, Callum Gapes, Andrea S. Griffin*]{style="font-size: 0.8em;"}

\

# [Material]{.smallcaps}

Multiple variables might influence the movement of migratory shorebirds within there foraging ground (ie. Hunter & Port Stephen estuaries), selecting different areas at different times whether to roost or forage. Those variables might vary across species and individuals, therefore their selected habitat. Within such an interdependent system, we try to suggest an approach that takes into account the tidal and circadian cycles as the main drivers to the habitat selection.

1.  **MOTUS telemetry technology**

<p style="font-size: 0.9em; color: gray;">

*Overview about the MOTUS array and its combination with Lotek VHF devices used within the study. How does this work?*

</p>

2.  **The local MOTUS array**

<p style="font-size: 0.9em; color: gray;">

*See how, where and where our array is set-up.*

</p>

3.  **The deployed tags**

<p style="font-size: 0.9em; color: gray;">

*Keep track on what tags are deployed over our migratory shorebirds.*

</p>

> Navigate the three different tabs below to get deeper into the [material]{.smallcaps} used.

::: panel-tabset
## Close tabs

## 1. MOTUS telemetry technology

To be continued ...

## 2. The local MOTUS array

To be continued ...

## 3. The deployed tags

To be continued ...
:::

# [Method]{.smallcaps}

This following aims to describe the analysis ran with VHF Motus data set that tries to access the complex articulation of habitat selection from migratory shorebirds across and within estuaries habitats. The analysis tackles several key points before properly studying migratory shorebirds spatial ecology as described as below:

1.  **Filtering the data**

<p style="font-size: 0.9em; color: gray;">

*Aims to clean out, simplify and format the raw data set from Motus network (.sql) to an easier table (.rds). Remove false positive signals, clarify ambiguous detection and add useful variables (ie. tide cycle, circadian cycle, etc.).*

</p>

2.  **Relative site usage**

<p style="font-size: 0.9em; color: gray;">

*The local deployed Motus array is composed of 8 receivers that suppose to 7/24 'listen' the tag signals, sent by the devices settled on birds, potentially within their range of reception. However, all the receivers did not 'listen' the same amount of time for multiple reasons (ie. date of networked, temporary out of servicing, etc.). Hence, the temporal coverage is unequal across the receivers meaning the survey effort -7/24 'listening'- is not comparable. This chapter access this bias by suggesting a site usage from bird, relative to the station effort survey.*

</p>

3.  **Motus array suitability**

<p style="font-size: 0.9em; color: gray;">

*All the bird we are focused on do not share the same land-use. That means the built-up array can not be suitable at the best for every species since the receivers are permanently settled at specific sites across the estuaries. Among all the deployed tags, which ones are the best surveyed within the local Motus array?*

</p>

4.  **Habitat selection**

<p style="font-size: 0.9em; color: gray;">

*Descriptive & inferential analysis are conducted.* *First is visually compared the spatial occupancy from birds and over our Motus array depending two main factors: tide and circadian cycles.*

</p>

> Navigate the four different tabs below to get deeper into the [method]{.smallcaps} used.

:::: panel-tabset
## Close tabs

## 1. Filtering the data

### [Contents]{style="color: grey; font-size: smaller;"}

-   [Packages](#packages)
-   [Download the data from Motus network](#download-the-data-from-motus-network)
-   [Filter the data](#filter-the-data)
-   [Adding variables](#adding-variables)

### Packages {#packages}

```{r 1 packages, message = FALSE, warning = FALSE, eval = TRUE, echo = TRUE}

# install.packages("motus", 
#                  repos = c(birdscanada = 'https://birdscanada.r-universe.dev',
#                            CRAN = 'https://cloud.r-project.org'))
library(motus)
library(dplyr)
library(here)
library(DBI)
library(RSQLite)
library(forcats) 
library(lubridate)
library(bioRad) 
library(purrr) 
```

### Download the data from Motus network {#download-the-data-from-motus-network}

```{r 1 settings, message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE}

# Global settings
setwd(dirname(rstudioapi::getSourceEditorContext()$path)) 
Sys.setenv(TZ="UTC") 
proj.num <- 294  
motusLogout()
```

Set general settings first as working directory, Time Zone, Motus project number (ie. ours is 294). And make sure the environment is free from any connection to any other networked project or from any other Motus user account (avoid undesired mistakes).

```{r 1 download data, message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE}
sql.motus <- tagme(projRecv = proj.num,
                   new = FALSE, # TRUE overwrites existing (large data takes a while) FALSE used to update existing file
                   update = TRUE,
                   dir = here("data"))

metadata(sql.motus, proj.num)

sql.motus <- dbConnect(SQLite(), here::here("data", "project-294.motus"))

df.alltags <- tbl(sql.motus, "alltags") %>%
  dplyr::collect() %>%
  as.data.frame()
```

`motus::tagme()`gets the data from the online Motus network, the ones part of your project (ie. ID = 294) only. Make sure you set your own directory. A file of type .sql is automatically created.

`motus::metadata()` downloads the metadata from the online Motus network (receiver information and more).

`motus::dbConnect()` links your .sql to your environment. Avoid to use high memory as .sql is a *lazy* table and not yet hardly written on your hardware.

`motus::tbl()` extracts all the tags fat into your environment. Then formatted into a classic dataframe.

### Filter the data {#filter-the-data}

```{r 1 filter the data, message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE}
# Wrong tags
df.alltags <- df.alltags %>% 
  #   filter(motusTagID == c("81121", "60470")) %>%

# Wrong receivers
   filter(!is.na(recvDeployLat),
          recvDeployName != c("Throsby Creek Test Site"),
          recv != c("SG-62A5RPI36710") ) # test_station
  
# False positive
df.alltags <- df.alltags %>% 
  filter(motusFilter == 1, # 0 is invalid data
         runLen >= 3) # value to be further thought

# Ambiguous 
clarify(sql.motus)
```

Here are cleaned-out the data recorded from undesired tags and/or receivers.

Then the False Positive signals are put away (ie. noise coming from external device, or any kind of external radio activity happening within the area) thanks to pre-made Motus filter and \`runLen\` threshold value, here set at 3. This value defines the number of bursts recorded from one tag and received at once, too low amount of burst are suspected for not being True Positif so not reliable.

`motus::clarify()` checks whether the combination of two tag signals emitting at the mean time could generate the detection of a third tag signal, which would be a False Positive on its own, but also hidding the two True Positives. If the table generated by this code has rows resulting, please go to [Motus documentation.](https://motuswts.github.io/motus/articles/05-data-cleaning.html).

### Adding variables {#adding-variables}

```{r 1 Tide data, message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE}
# Load tide (Callum work 01_import_tide_data.R)
tidalCurve <- readRDS(here::here("10_data", "tides", "tidalCurve.rds"))
tideData <- readRDS(here::here("10_data", "tides", "tideData.rds"))
tidalCurveFunc <- splinefun(tideData$tideDateTimeAus, tideData$tideHeight, method = "natural")
get.tideIndex <- function(time){ return(which.min(abs(tideData$tideDateTimeAus-time)))}
```

Note that the raw Tide data-set is not provided here. Since it has been manually extracted from [New South Wales government resources](https://www.nsw.gov.au/sites/default/files/2024-06/tide-tables-24-25-nsw.pdf) and formatted by [Callum Gapes](https://www.linkedin.com/in/callum-g/?originalSubdomain=au) through this [method](https://uoneduau-my.sharepoint.com/:u:/g/personal/c3541851_uon_edu_au/EWOk05RFs1xJhAWLBddbf-kB_H2xRuz0XPUWtPVfcju_-Q?e=P6WzOT).

```{r 1 variable, message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE}
df.alltags <- df.alltags %>% 
  
# Time
  mutate(time = as_datetime(ts),
         timeAus = as_datetime(ts, tz = "Australia/Sydney"),
         dateAus = as_date(timeAus),
         year = year(time), # extract year from time
         doy = yday(time)) %>%
  
# Sunrise/set
  sunRiseSet(lat = "recvDeployLat", 
             lon = "recvDeployLon", 
             ts = "ts") %>% 
  mutate(sunriseNewc = sunrise(dateAus, 151.7833, -32.9167, elev = -0.268, tz = "Australia/Sydney", force_tz = TRUE),
         sunsetNewc = sunset(dateAus, 151.7833, -32.9167, elev = -0.268, tz = "Australia/Sydney", force_tz = TRUE)) %>%
  
# Positive signal strength (min. = 0) for plotting
  mutate(sigPositive = sig + abs(min(sig))) %>%
  
# As Factor
  mutate(motusTagID = as.factor(motusTagID))

# Tide
df.alltags <- df.alltags %>% 
  mutate(tideHeight = tidalCurveFunc(timeAus),
         tideIndex = map_dbl(timeAus, get.tideIndex))

tide_values <- tideData[df.alltags$tideIndex, 
                        c("tideDateTimeAus",
                          "high_low",
                          "day_night",
                          "tideCategory",
                          "tideID",
                          "tideHeight")]

df.alltags <- df.alltags %>%
  mutate(tideDateTimeAus = tide_values$tideDateTimeAus,
         tideHighLow = as_factor(tide_values$high_low),
         tideDiel = as_factor(tide_values$day_night),
         tideCategory = as_factor(tide_values$tideCategory),
         tideCategoryHeight = tide_values$tideHeight,
         tideID = as_factor(tide_values$tideID),
         tideTimeDiff = abs(difftime(timeAus, tideDateTimeAus, units = "hours")) ) # Time diff btw detect. & nearest tide pts

```

Several variables are added as some are formatted and/or renamed for the aims of the study.

```{r 1 recv, message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE}
# Get summary
df.recvDeps <- tbl(sql.motus, "recvDeps") %>% 
  collect() %>% 
  as.data.frame()

# Rename stations
station_rename <- list(
   "Barry_Fullerton_cove"  = "Fullerton Entrance",
   "North Swann Pond"      = "Swan Pond" ,
   "Ramsar Road Floodgate" = "Ramsar Road",
   "Milham's Pond"         = "Milhams Pond")
df.recvDeps <- df.recvDeps %>% 
   mutate(recvDeployName = recode(stationName,
                            !!!station_rename))
df.alltags <- df.alltags %>% 
  mutate(recvDeployName = recode(recvDeployName,
                                 !!!station_rename))
```

Rename here the receiver of your array for a better comprehension.

```{r 1 save .rds, message = FALSE, warning = FALSE, eval = FALSE, echo = TRUE}
saveRDS(df.alltags, here::here("data", paste0(Sys.Date(), "-data", ".rds" )))
saveRDS(df.recvDeps, here::here("data", paste0(Sys.Date(), "-recv-info", ".rds" )))
```

Save your now formatted data, ready for use. Automatically named with the date of your saving.

## 2. Relative site usage

## 3. Motus array suitability

## 4. Habitat selection

### [Contents]{style="color: grey; font-size: smaller;"}

-   [Packages](#packages2)
-   [Load the data](#load-the-data)
-   [Summarise the data](#summarise-the-data)
-   [Descriptive analysis](#descriptive-analysis)

### Packages

```{r 4 packages, message = FALSE, warning = FALSE, eval = TRUE}
library(motus)
library(dplyr)
library(here)
library(DBI)
library(RSQLite)
library(forcats) 
library(lubridate)
library(bioRad) 
library(purrr) 
library(sf)
library(lubridate)
library(rnaturalearth)
library(ggplot2)
```

### Load the data {#load-the-data}

Either load your own data generated across the previous chapters (.rds format), or download the data up to the-last document-update date provided here below:

Receiver information: <a href="#" onclick="return checkPassword('recv')">recv-info.rds</a><br> Bird data: <a href="#" onclick="return checkPassword('bird')">data.rds</a>

```{=html}
<script>
  function checkPassword(type) {
    var correctPasswords = {
      recv: "provence13",
      bird: "provence13"
    };
    var urlLinks = {
      recv: "https://uoneduau-my.sharepoint.com/:u:/g/personal/c3541851_uon_edu_au/ESvl2B4fx7FCsKMiEXrgnyABGycGYU0RV-PfOftI0wsTbA?e=OdR3ag",
      bird: "https://uoneduau-my.sharepoint.com/:u:/g/personal/c3541851_uon_edu_au/EbZbqh6eidVMksDUyl-J6UcBV8skhQXsjmL5f5nBw1pnwg?e=fJcygo"
    };
    var userPassword = prompt("Please enter the password to access the link:");
    if (userPassword === correctPasswords[type]) {
      window.open(urlLinks[type], '_blank');
    } else {
      alert("Incorrect password. Access denied.");
    }
    return false; // prevent default link behavior
  }
</script>
```

To properly load the data into your environment, make sure to set the working directory and the path as yours.

```{r 4 load data 1, eval = FALSE}
# Set WD
setwd(dirname(rstudioapi::getSourceEditorContext()$path)) 
# Birds
data_all <- readRDS(here::here("data", "recv.info.rds"))
# Receivers info
recv <- readRDS(here::here("data", "data.rds")) 
```

```{r 4 load data 2, echo = FALSE, eval = TRUE}

# Birds
data_all <- readRDS(
  tail(sort(list.files(
  here::here("10_data", "alltags", "motus.rds"),
  pattern = "-data\\.rds$", full.names = TRUE
  )), 1)) # pick up the most recent .rds file

# Receivers info
recv <- readRDS(
  tail(sort(list.files(
    here::here("10_data", "alltags", "motus.rds"),
    pattern = "-recv-info\\.rds$", full.names = TRUE
  )), 1)) 
```

### Summarise the data {#summarise-the-data}

For the aim of the analysis and generate readable figures, let's simplify the table to keep useful information only.

```{r 4 summarise data, echo = TRUE, eval = TRUE}
# Data summary (To be edited still...)
tagSummary1 <- data_all %>%
  group_by(motusTagID, recvDeployName) %>% 
  summarize(nDet = n(),
            nRecv = n_distinct(recvDeployName),
            timeMin = min(time),
            timeMax = max(time),
            totDay = length(unique(doy)), 
            species = first(speciesEN),
            .groups = "drop") %>%
  relocate(motusTagID, species)

tag_summary2 <- data_all %>%
  group_by(recvDeployName, tideCategory, speciesEN) %>%
  summarise(nTags = n_distinct(motusTagID), .groups = "drop")

# Selecting variables
data <- data_all %>%
  select(
    
  # Project ID
    recvProjID,
    tagProjID,
  # Tag
    motusTagID,
  # Time
    time,
    timeAus,
    year,
  # Receiver
    recv,
    recvDeployName,
  # Tide cycle
    tideHeight,
    tideCategory,
  # Circadian cycle
    sunriseNewc,
    sunsetNewc,
  # Signal strengh
    sigPositive,
  # Bird info
    speciesID, speciesEN, speciesFR, speciesSci,
    speciesGroup,
  # Comment
    tagDepComments
    )

# Color codes
species_colors <- c(
  "Bar-tailed Godwit"      = "#1b9e77",  
  "Far Eastern Curlew"     = "#d95f02",  
  "Masked Lapwing"         = "#7570b3", 
  "Pacific Golden-Plover"  = "#e7298a", 
  "Pied Stilt"             = "#66a61e", 
  "Red-necked Avocet"      = "#e6ab02"   
)
```

#### Descriptive analysis {#descriptive-analysis}

[Tide & circadian cycles](#tide-circadian-cycle)


##### Tide & circadian cycles {#tide-circadian-cycle}

-   Over the Motus array

    ```{r 4 Habitat selection across sites, echo = TRUE, eval = TRUE}
    tag_summary2 %>%
      mutate(tideCategory  = case_when(
      tideCategory == "Nocturnal_High" ~ "NH",
      tideCategory == "Diurnal_Low"    ~ "DL",
      tideCategory == "Diurnal_High"   ~ "DH",
      tideCategory == "Nocturnal_Low"  ~ "NL",
      TRUE ~ tideCategory )) %>%
      ggplot(aes(x = as.factor(tideCategory), y = nTags, fill = speciesEN)) +
      geom_col(position = "stack") +   
      facet_wrap(~ recvDeployName) +   
      theme_bw() +
      labs(x = "Tide Category", y = "Number of individual detected", fill = "Species") +
      scale_fill_manual(values = species_colors)
    ```

-   Over each Motus stations

::: panel-tabset
```{r 4 Habitat selection across sites 2, echo = FALSE, eval = TRUE, results = 'asis'}
plots <- tag_summary2 %>%
      mutate(tideCategory  = case_when(
      tideCategory == "Nocturnal_High" ~ "NH",
      tideCategory == "Diurnal_Low"    ~ "DL",
      tideCategory == "Diurnal_High"   ~ "DH",
      tideCategory == "Nocturnal_Low"  ~ "NL",
      TRUE ~ tideCategory )) %>%
      split(.$recvDeployName) %>%
      map(~ ggplot(.x, aes(x = tideCategory, y = nTags, fill = speciesEN)) +
            geom_col(position = "stack") +
            facet_wrap(~ speciesEN) +
            scale_fill_manual(values = species_colors) +
            theme_bw() +
            labs(x = "Tide Category",
                 y = "Number of individuals detected",
                 fill = "Species"))
```

```{r 4 Plot 2, echo = FALSE, eval = TRUE, results = 'asis'}
  for (recvDeployName in names(plots)) {
    cat("\n\n## ", recvDeployName, "\n\n") 
    print(plots[[recvDeployName]]) 
    flush.console() #force immediate output
  }
```
:::

Reminder: Using the term *habitat selection* report in fact to Motus receiver area coverage selection.
::::

***Availability***

*The analysis are ran through R version 4.5.1 (2025-06-13 ucrt) -- "Great Square Root" and the platform: x86_64-w64-mingw32/x64.*
