---
title: "Shorebird_Monitoring_NSW_Callum"
author: "Callum Gapes"
date: "2023-08-14"
output: html_document
---
## Setup RStudio and Import Motus Data

```{r Setup, results = "hide"}
rm(list = ls()) # Clear environment
Sys.setenv(TZ="UTC") # Set timezone to UTC
setwd("C:/Users/callu/My Drive/Documents/UoN/SCIE3500/Shorebird Monitoring NSW") # Set working directory
```

## Load Required Packages

```{r Packages, results = "hide", message = FALSE}
library(lubridate)
library(motus)
library(motusData)
library(dplyr)
library(DBI)
library(RSQLite)
library(lutz)
library(ggplot2)
library(crayon)
library(tidyverse)
library(rnaturalearth)
library(ggmap)
library(sf) # Spatial vector data
library(suntools) # Sunrise/set info
library(ggnewscale)
library(gginnards) # Moving plot layers (tide on bottom)

motusLogout()
```

## Plot Setup (All Plots)
```{r Plot Setup}
plotSun <- 0
plotTide <- 0
plotTidalCurve <- 1
completeTimeRange <- 0

sunriseColour = "orange"
sunsetColour = "purple"

highTideColour = "blue"
lowTideColour = "#61caff"
tideColour = "#61caff"

vlineAlpha <- 0.3
vlineWidth <- 1
pointSize <- 0.8
```

```{r Custom Functions}

# Get species common name from tag ID
get.species <- function(tagID) {
  return(unique(filter(df.alltags,motusTagID==tagID)$speciesEN))
}

save.plot <- function(title = paste(p$labels$title," ",format(currentTime,"%Y-%m-%d %H-%M-%S"),".png",sep = "")) {
  currentTime <- with_tz(Sys.time(),"Australia/Sydney")
  #title <- paste(p$labels$title," ",format(currentTime,"%Y-%m-%d %H-%M-%S"),".png",sep = "")
  if (title != paste(p$labels$title," ",format(currentTime,"%Y-%m-%d %H-%M-%S"),".png",sep = "")){
    title <- paste(title,".png",sep = "")
  }
  ggsave(title,plot = p, path = "Plots")
  return()
}

plot.addSunriseSet <- function(p){
  p <- p + ggnewscale::new_scale_colour() +
  geom_vline(aes(xintercept=sunrise,colour="sunriseColour"),alpha = 0.3,linewidth = 1) +
  geom_vline(aes(xintercept=sunset,colour="sunsetColour"), alpha = 0.3,linewidth = 1) +
  scale_color_manual(name = "Sun",
        labels = c("Sunrise","Sunset"),
        values = c(sunriseColour,sunsetColour),
        guide = guide_legend(order = 2))
  return(p)
}

plot.addTideHighLow <- function(p){
  p <- p + ggnewscale::new_scale_colour() +
  geom_vline(data = filter(tideData, high_low == "High"), aes(xintercept = tideDateTimeAus, colour = high_low),
                                                              alpha = 0.3,
                                                              linewidth = 1) +
  geom_vline(data = filter(tideData, high_low == "Low"), aes(xintercept = tideDateTimeAus, colour = high_low),
                                                             alpha = 0.3, linewidth = 1) +
  scale_color_manual(name = "Tides",
        labels = c("High Tide","Low Tide"),
        values = c(highTideColour,lowTideColour),
        guide = guide_legend(order = 3))
  return(p)
}

plot.addTidalCurve <- function(p){
  # Filter tidal curve by time range
  tmpTide <- filter(tidalCurve, time >= timeStart, time <= timeEnd)
  
  # Scaling factor to plot tide on the same axes as signal strength
  scaleCoeff <- (max(tmpTide$height)-min(tmpTide$height))/(max(tmp$sigPositive)-min(tmp$sigPositive))
  
  # Vertical shift amount to centre tidal curve on plot
  verticalShift <- ((max(tmpTide$height/scaleCoeff)+min(tmpTide$height/scaleCoeff))/2) - ((max(tmp$sigPositive)+min(tmp$sigPositive))/2)

  # Plot tidal curve
  p <- p + geom_line(data = tmpTide, aes(x = time, y = height/scaleCoeff-verticalShift) ,colour = tideColour, linewidth = 3, alpha = 0.3)  +
  theme(
    axis.title.y.right = element_text(size = 13)
  ) 
  
  # Add second axis for tide height
  p <- p +  scale_y_continuous(
    sec.axis = sec_axis(~.*scaleCoeff+verticalShift*scaleCoeff, name = "Tide Height (m)")
)
  return(p)
}

```


```{r Recent Detections}
mostRecentDetections <- function(){
  for (i in 1:receiverCount) {
    receiverInfo$recentDetection[i] <- df.alltags %>% filter(recvDeployName == receiverNames[i]) %>% select(ts) %>% max()
    receiverInfo$recentDetectionFiltered[i] <- df.alltags %>% filter(recvDeployName == receiverNames[i],motusFilter == 1) %>% select(ts) %>% max()
  }
  
  receiverInfo <- mutate(receiverInfo, recentDetectionAus = recentDetection %>% as_datetime(tz = "Australia/Sydney"))
  receiverInfo <- mutate(receiverInfo, recentDetectionFilteredAus = recentDetectionFiltered %>% as_datetime(tz = "Australia/Sydney"))
  receiverInfo <- receiverInfo %>% select(name, recentDetectionAus, recentDetectionFilteredAus)
  
  print(receiverInfo)
  
  return(invisible(NULL))
}

```


Define tags and receivers of interest
```{r Tags and Receivers}
receiverNames <- c("Milham's Pond","Barry_Fullerton_cove", "North Swann Pond", "Ramsar Road Floodgate")
#receiverNames <- df.alltags$recvDeployName %>% unique()
tagIDs <- c(60472, 43423, 43289, 43298, 43301, 43302, 43304, 43305, 43306, 60473, 60577, 60578)
tagIDs <- append(tagIDs, c(43296, 43300, 43303, 60471)) # Add new birds tagged at Stockton 29/08/2023

receiverCount <- length(receiverNames)
receiverInfo <- as.data.frame(receiverNames)
colnames(receiverInfo) <- c("name")
```


Import Motus Data, Convert to Data Frame and Filter Receivers
```{r}
## Import Motus Data From File
project294.motus <- dbConnect(SQLite(), "project-294.motus")

# Extract "alltags" table, which contains the detection data
tbl.alltags <- tbl(project294.motus, "alltags")

# Filter by receivers and tags of interest
tbl.alltags <- tbl.alltags %>% 
  filter(recvDeployName %in% receiverNames, motusTagID %in% tagIDs)

# Filter only by tags of interest
#tbl.alltags <- tbl.alltags %>% 
#  filter(motusTagID %in% tagIDs)

# Convert to data frame for manual data exploration
df.alltags <- tbl.alltags %>% collect() %>%  as.data.frame()

## Format timestamps

# Add datetime time stamp column - in both UTC (default) and Australian time
df.alltags <- df.alltags %>% mutate(time = as_datetime(ts))
df.alltags <- df.alltags %>% mutate(timeAus = as_datetime(ts, tz = "Australia/Sydney"))

# Add column for date (without time)
df.alltags$dateAus <- df.alltags$timeAus %>% as_date()

## Add modified signal strength for plotting

# Scale signal strength to positive (where minimum = 0) to make plotting easier
df.alltags <- df.alltags %>% mutate(sigPositive = sig + abs(min(sig)))
```


## Import Sunset/Sunrise Data
```{r Additional Data}
df.alltags <- sunRiseSet(df.alltags, lat = "recvDeployLat", lon = "recvDeployLon", ts = "ts")
```
## Import NSW Tides from Spreadsheet and Generate Tidal Curve
```{r Import Tide Data}
tideData <- read.csv("TideDataNewcastle2023.csv")

# Format date and datetime columns
tideData$tideDateTimeAus <- ymd_hms(tideData$tideDateTime, tz = "Australia/Sydney")

# Generate tidal curve
# NOTE: Spline fits tide data quite well, EXCEPT for the second and second last values
tidalCurve <- as_data_frame(spline(tideData$tideDateTimeAus, tideData$tideHeight, n = 1000*length(tideData$tideDateTimeAus), method = "natural"))
tidalCurve$x <- as_datetime(tidalCurve$x, tz = "Australia/Sydney")

colnames(tidalCurve) <- c("time","height")
```




## SETUP: RUN ALL SECTIONS ABOVE HERE (Ctrl + Alt + Shift + P)

## Download Motus Detection Data NOTE
ATTEMPTED TO DOWNLOAD DETECTIONS: GOT ERROR
```{r, eval = FALSE}
# Specify Motus project number
proj.num <- 294

## Download project detections across motus network
# Change new = TRUE if running for the first time (to create the .motus SQLite databse on your computer)
sql.motus <- tagme(projRecv = proj.num, new = FALSE, update = TRUE)

# download detections on my project's receivers only
sql.motus <-tagme(proj.num, new=FALSE, update = TRUE)
metadata(sql.motus,proj.num)
tbl.recvDeps <- tbl(sql.motus, "recvDeps")

df.serno <- tbl.recvDeps %>%
  filter(projectID == proj.num) %>%
  select(serno) %>%
  distinct() %>%
  collect() %>% as.data.frame()

#loop through each receiver 
for(row in 1:nrow(df.serno)) {
  tagme(df.serno[row, "serno"], new = FALSE, update = TRUE)
}
```



## Explore Structure of Motus Data
```{r, eval = FALSE, echo = FALSE}
# List of tables within .motus file
dbListTables(project294.motus)

# Extract "alltags" table, which contains the detection data
tbl.alltags <- tbl(project294.motus, "alltags")

# Convert to data frame for manual data exploration
df.alltags <- tbl.alltags %>% collect() %>%  as.data.frame()

names(df.alltags) # field names
str(df.alltags) # variable types

```



## Plot detections for single tag
```{r Plot: Single Tag}

# Choose tag to plot
currentTagID <- 60472

# Set start and end time for plot
if (completeTimeRange) {
  # Get time interval from first and last detections
  timeStart <- min(df.alltags$timeAus)
  timeEnd <- max(df.alltags$timeAus)
} else{
  # Set time interval manually
  timeStart <- ymd_hms("2023-04-16 09:00:00",tz = "Australia/Sydney")
  timeEnd <- ymd_hms("2023-04-24 21:00:00",tz = "Australia/Sydney")
}

# Filter by tag and time interval
tmp <- filter(df.alltags, motusTagID == currentTagID, timeAus >= timeStart, timeAus <= timeEnd)

# Create plot
p <- tmp %>% 
  ggplot(aes(timeAus, sigPositive)) + # x-axis = signal strength, y-axis = time
  
  # Plot points
  geom_point(aes(colour = recvDeployName), 
             size = pointSize, 
             shape = "circle") + 
  
  ## Automatic Legend
  #labs(colour = "Receiver") + # Legend title
  
  ## Manually labelled legend
  scale_colour_discrete(name = "Receiver",
                        labels = c("Receiver 1", "Receiver 2")) +
  
  guides(colour = guide_legend(order = 1)) + # Make this legend appear first
  
  # Axis titles
  ylab("Signal Strength") + 
  xlab("Date-Time") + 
  
  ## Automatic Plot Title
  #ggtitle(paste("Detections for Tag",currentTagID)) + theme(plot.title = element_text(hjust = 0.5)) 
  
  ## Custom Plot Title
  ggtitle("Tag 60472 - Eastern Curlew") + theme(plot.title = element_text(hjust = 0.5)) 

# Add sunset and sunrise
if (plotSun) {
  p <- plot.addSunriseSet(p)
}

# Add high and low tides
if (plotTide) {
  p <- plot.addTideHighLow(p)
}

# Add tidal curve
if (plotTidalCurve) {
  p <- plot.addTidalCurve(p)
}

p
  
```
## Plot detections for all tags, grid by tag, grouped by receiver
```{r Plot: Grid - Tag, Colour - Receiver}

# Set start and end time for plot
if (completeTimeRange) {
  # Get time interval from first and last detections
  timeStart <- min(df.alltags$timeAus)
  timeEnd <- max(df.alltags$timeAus)
} else{
  # Set time interval manually
  timeStart <- ymd_hms("2023-04-10 09:00:00",tz = "Australia/Sydney")
  timeEnd <- ymd_hms("2023-04-12 09:00:00",tz = "Australia/Sydney")
}

# Filter by time range
tmp <- filter(df.alltags, timeAus >= timeStart, timeAus <= timeEnd)   
  
p <- tmp %>% 
  ggplot(aes(timeAus, sigPositive)) +
  
  # Plot points
  geom_point(aes(colour = recvDeployName), size=0.3) +
  facet_grid(vars(motusTagID)) +
  labs(colour = "Receiver") + # Legend title
  guides(colour = guide_legend(order = 1)) +

  ylab("Signal Strength") + xlab("Date-Time") + # Axis titles
  ggtitle("Detections by all Receivers ") + theme(plot.title = element_text(hjust = 0.5)) # Plot title

# Add sunset and sunrise
if (plotSun) {
  p <- plot.addSunriseSet(p)
}

# Add high and low tides
if (plotTide) {
  p <- plot.addTideHighLow(p)
}

# Add tidal curve
if (plotTidalCurve) {
  p <- plot.addTidalCurve(p)
}

p

```
## Plot detections for single receiver
```{r Plot: Single Receiver}
# Choose receiver to plot
currentReceiver <- "Barry_Fullerton_cove"

# Set start and end time for plot
if (completeTimeRange) {
  # Get time interval from first and last detections
  timeStart <- min(df.alltags$timeAus)
  timeEnd <- max(df.alltags$timeAus)
} else{
  # Set time interval manually
  timeStart <- ymd_hms("2023-08-29 17:00:00",tz = "Australia/Sydney")
  timeEnd <- ymd_hms("2023-08-30 12:00:00", tz = "Australia/Sydney")
}

# Filter by receiver and time interval
tmp <- filter(df.alltags, recvDeployName == currentReceiver, timeAus >= timeStart, timeAus <= timeEnd)

p <- tmp %>% 
  ggplot(aes(x = timeAus, y = sigPositive)) + # x-axis = signal strength, y-axis = time
  
  geom_point(aes(colour = as.character(motusTagID)),
                 shape = "circle",
                 size = 0.5) + # plot points
  labs(colour = "Motus Tag ID", ) + # Legend title
  guides(colour = guide_legend(order = 1)) +

  ylab("Signal Strength") + xlab("Date-Time") + # Axis titles
  theme(axis.title.y = element_text(size = 13)) + # y-axis title format
  
  ## Automatic title 
  #ggtitle(paste("Detections at",currentReceiver)) + theme(plot.title = element_text(hjust = 0.5)) # Plot title
  
  ## Custom title
  ggtitle("Pied Stilt Detections") + theme(plot.title = element_text(hjust = 0.5)) # Plot title

  
# Add sunset and sunrise
if (plotSun) {
  p <- plot.addSunriseSet(p)
}

# Add high and low tides
if (plotTide) {
  p <- plot.addTideHighLow(p)
}

# Add tidal curve
if (plotTidalCurve) {
  p <- plot.addTidalCurve(p)
}

  
p
  
```
## Plot detections for all tags, grid by receiver, colour by species
```{r Plot" Grid - Receiver, Colour - Species}
# Set start and end time for plot
if (completeTimeRange) {
  # Get time interval from first and last detections
  timeStart <- min(df.alltags$timeAus)
  timeEnd <- max(df.alltags$timeAus)
} else{
  # Set time interval manually
  timeStart <- ymd_hms("2023-04-10 09:00:00",tz = "Australia/Sydney")
  timeEnd <- ymd_hms("2023-04-12 09:00:00",tz = "Australia/Sydney")
}

tmp <- filter(df.alltags, timeAus >= timeStart, timeAus <= timeEnd)
  
p <- tmp %>% 
  ggplot(aes(timeAus, sigPositive)) +
  
  geom_point(aes(colour = speciesEN), size=0.8) +
  facet_grid(vars(recvDeployName)) +
  labs(colour = "Species") + # Legend title
  guides(colour = guide_legend(order = 1)) +
  
  ylab("Signal Strength") + xlab("Date Time") + # Axis titles
  ggtitle("Detections by all Receivers ") + theme(plot.title = element_text(hjust = 0.5)) # Plot title

# Add sunset and sunrise
if (plotSun) {
  p <- plot.addSunriseSet(p)
}


# Add high and low tides
if (plotTide) {
  p <- plot.addTideHighLow(p)
}

# Add tidal curve
if (plotTidalCurve) {
  p <- plot.addTidalCurve(p)
}

p
```

## Plot detections for all tags, grid by receiver, colour by tag
```{r Plot: Grid - Receiver, Colour - Tag}
# Set start and end time for plot
if (completeTimeRange) {
  # Get time interval from first and last detections
  timeStart <- min(df.alltags$timeAus)
  timeEnd <- max(df.alltags$timeAus)
} else{
  # Set time interval manually
  timeStart <- ymd_hms("2023-04-15 09:00:00",tz = "Australia/Sydney")
  timeEnd <- ymd_hms("2023-04-20 09:00:00",tz = "Australia/Sydney")
}

tmp <- filter(df.alltags, timeAus >= timeStart, timeAus <= timeEnd)

p <- tmp %>% 
  ggplot(aes(timeAus, sigPositive)) +
  
  geom_point(aes(colour = as.character(motusTagID)), size=0.3) +
  facet_grid(vars(recvDeployName)) +
  labs(colour = "Motus Tag ID") + # Legend title
  guides(colour = guide_legend(order = 1)) +

  ylab("Signal Strength") + xlab("Date Time") + # Axis titles
  ggtitle("Detections by all Receivers ") + theme(plot.title = element_text(hjust = 0.5)) # Plot title

# Add sunset and sunrise
if (plotSun) {
  p <- plot.addSunriseSet(p)
}

# Add high and low tides
if (plotTide) {
  p <- plot.addTideHighLow(p)
}

# Add tidal curve
if (plotTidalCurve) {
  p <- plot.addTidalCurve(p)
}

p
```

## Plot detections for all tags, grid by species, colour by receiver
```{r Plot: Grid - Species, Colour - Receiver}
# Set start and end time for plot
if (completeTimeRange) {
  # Get time interval from first and last detections
  timeStart <- min(df.alltags$timeAus)
  timeEnd <- max(df.alltags$timeAus)
} else{
  # Set time interval manually
  timeStart <- ymd_hms("2023-04-12 09:00:00",tz = "Australia/Sydney")
  timeEnd <- ymd_hms("2023-04-15 09:00:00",tz = "Australia/Sydney")
}

tmp <- filter(df.alltags, timeAus >= timeStart, timeAus <= timeEnd)
  
p <- tmp %>% 
  ggplot(aes(timeAus, sigPositive)) +
  
  geom_point(aes(colour = recvDeployName), size=0.8) +
  facet_grid(vars(speciesEN)) +
  labs(colour = "Receiver") + # Legend title
  guides(colour = guide_legend(order = 1)) + 
  scale_colour_discrete(name = "Receiver",
                        labels = c("Receiver 1", "Receiver 2","Receiver 3","Receiver 4")) +
  
  ylab("Signal Strength") + xlab("Date Time") + # Axis titles
  ggtitle("Detections of Tagged Species ") + theme(plot.title = element_text(hjust = 0.5)) # Plot title

# Add sunset and sunrise
if (plotSun) {
  p <- plot.addSunriseSet(p)
}


# Add high and low tides
if (plotTide) {
  p <- plot.addTideHighLow(p)
}

# Add tidal curve
if (plotTidalCurve) {
  p <- plot.addTidalCurve(p)
}

p
```

```{r Experimenting with Tide Data}
# Select date range
timeStart <- ymd_hms("2023-04-10 09:00:00",tz = "Australia/Sydney")
timeEnd <- ymd_hms("2023-04-20 09:00:00",tz = "Australia/Sydney")
  
# Filter date range
tmp <- filter(tideData, tideDateTimeAus >= timeStart, tideDateTimeAus <= timeEnd)

# Scatter plot
#ggplot(tmp, aes(x = tideDateTimeAus, y = tideHeight)) + geom_point()

# Sinusoidal regression
#sinlist <- sinreg(as.numeric(x), tmp$tideHeight, plot = TRUE)

# Test for single period
#high <- tmp$tideHeight[1]
#low <- tmp$tideHeight[2]
#highTime <- tmp$tideDateTimeAus[1]
#lowTime <- tmp$tideDateTimeAus[2]

#x <- seq(lowTime,highTime,length.out=1000)


# Filter tidalCurve by time period
tmpTide <- filter(tidalCurve, time >= timeStart, time <= timeEnd)

#
#out = spline(tmp$tideDateTimeAus, tmp$tideHeight, n = 1000*length(tmp$tideDateTimeAus), method = "natural")
#out <- as_data_frame(out)

p <- ggplot(tmpTide, aes(x=x,y=y))+geom_line()+geom_point(data = tmp, aes(x=tideDateTimeAus, y = tideHeight))


```

## TO DO - FUNCTION FOR FINDING MOST RECENT DETECTION AT EACH SITE


## Data Summarising (Motus Docs Article 03)
```{r}
file.name = project294.motus

# List tables within motus sqlite file
dbListTables(project294.motus)

# List of variables (fields) within a table
dbListFields(file.name, "runs")

# Find number of detections of each tag by each receiver
df.detectSum <- df.alltags %>%
  count(motusTagID, recv) %>% 
  collect() %>% 
  as.data.frame()

# Add species name
for (i in 1:nrow(df.detectSum)){
   df.detectSum$species[i] <- get.species(df.detectSum$motusTagID[i])
}
```

## Deployment Information (Motus Docs Article 04)
```{r}
## Check number of tags registered to Motus project
tbl.tags <- tbl(project294.motus, "tags") 
df.tags <- tbl.tags %>%
  filter(projectID == 294) %>%
  collect() %>%
  as.data.frame()

nrow(df.tags) # number of registered tags in the database

## Check number of registered tags that were deployed
tbl.tagDeps <- tbl(project294.motus, "tagDeps") 

df.tagDeps <- tbl.tagDeps %>%
  filter(projectID == 294) %>%
  collect() %>%
  as.data.frame() %>% # once in df format, can format dates with lubridate
  mutate(timeStart = as_datetime(tsStart),
         timeEnd = as_datetime(tsEnd)) 

anti_join(df.tags, df.tagDeps, by = "tagID") 

## Check if any tags have multiple deployments (which would make motusTagID unsuitable as a unique identifier)
df.alltags %>%
  select(motusTagID, tagDeployID) %>%
  filter(!(is.na(tagDeployID))) %>% # remove NA tagDeployIDs
  distinct() %>%
  group_by(motusTagID) %>%
  mutate(n = n()) %>%
  filter(n > 1)
```
# Tag deployment locations
```{r}
#world <- ne_countries(scale = "medium", returnclass = "sf") 
#lakes <- ne_download(scale = "medium", type = 'lakes', category = 'physical',
 #                    returnclass = "sf", destdir = "map-data") # only need this first time downloading
lakes <- ne_load(type = "lakes", scale = "medium", category = 'physical',
                 returnclass = "sf",
                 destdir = "map-data") # use this if already downloaded shapefiles

xmin <- min(df.tagDeps$longitude, na.rm = TRUE) - 5
xmax <- max(df.tagDeps$longitude, na.rm = TRUE) + 5
ymin <- min(df.tagDeps$latitude, na.rm = TRUE) - 5
ymax <- max(df.tagDeps$latitude, na.rm = TRUE) + 5
                
# map using ggplot and sf
ggplot(data = world) + 
  geom_sf(colour = NA) +
  geom_sf(data = lakes, colour = NA, fill = "white") +
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax), expand = FALSE) +
  labs(x = "", y = "") + 
  theme_bw() + 
  geom_point(data = filter(df.tagDeps, projectID == 294), 
             aes(longitude, latitude), size = 2, shape = 19, colour = "red")
```
## Number of project receiver deployments
```{r}
tbl.recvDeps <- tbl(project294.motus, "recvDeps") 

df.projRecvs <- tbl.recvDeps %>%
  filter(projectID == 294) %>%
  collect() %>%
  as.data.frame() %>%
  mutate(timeStart = as_datetime(tsStart),
         timeEnd = as_datetime(tsEnd))

df.projRecvs %>%
  mutate(dateStart = date(timeStart)) %>% 
  select(-serno,-fixtureType, -macAddress, -timeStart, -timeEnd, -elevation, 
         -projectID, -status, -receiverType, -siteName) %>%
  arrange(deviceID, latitude, dateStart)

## Filter by names
df.projRecvs <- df.projRecvs %>%
  filter(name %in% receiverInfo$name)

## Plot deployment periods

# put data in long format to simplify plotting (or use geom_segment)
df.projRecvs.long <- df.projRecvs %>%
  select(deviceID, name, timeStart, timeEnd) %>% 
  gather(when, time, c(timeStart, timeEnd)) %>%
  # fake end date:
  mutate(time = if_else(is.na(time), max(time, na.rm = TRUE) + duration(6, "month"), time)) 

ggplot(data = df.projRecvs.long, 
       aes(x = time, y = as.factor(deviceID), colour = as.factor(name))) +
  theme(legend.position = "none") +
  geom_line(lwd = 3) + 
  # instead, centre to the right
  geom_text(data = filter(df.projRecvs.long, when == "timeStart"), 
            aes(label = name), hjust = "left", nudge_y = 0.2, size = 3, angle = 45) +
  scale_color_viridis_d() +
  theme_bw() +
  #labs(x = "Year", y = "Receiver ID")
  labs(x = "Year", y = "Name")
```
## Check receiver and antenna metadata
```{r}
# Load all receiver metadata
df.recvDeps <- tbl.recvDeps %>%
  collect() %>%
  as.data.frame() %>%
  mutate(timeStart = as_datetime(tsStart),
         timeEnd = as_datetime(tsEnd))

# antenna metadata for ALL Motus antenna deployments; 
# to simplify, keep only the variables of interest.
tbl.antDeps <- tbl(project294.motus, "antDeps") 

df.antDeps <- tbl.antDeps %>%
  select(deployID, port, antennaType, bearing, heightMeters) %>%
  collect() %>%
  as.data.frame()

# receiver deployments; select variables of interest
df.recvDeps <- df.recvDeps %>%
  select(deployID, receiverType, deviceID, name, latitude, longitude, 
         isMobile, timeStart, timeEnd, projectID, elevation) 

df.stationDeps <- left_join(df.recvDeps, df.antDeps, by = "deployID")
#df.stationDeps <- filter(df.stationDeps, projectID == 294)

# Filter by specific list of receivers (based on name)
df.stationDeps <- filter(df.stationDeps, name %in% receiverInfo$name)

summary(df.stationDeps)

# Filter deployments with missing bearing
df.stationDeps %>%
  filter(is.na(bearing)) %>%
  select(-elevation, -deviceID, -timeEnd)

# 
```

## Data Cleaning (Motus Docs Article 05)
```{r}
## Check how many detections (hits) have motusFilter = 0 (meaning they are likely 'dubious')
df.alltags %>% filter(motusFilter == 0) %>% count()


```

## Occurence by day by tag
```{r}
dateStart <- ymd("2023-01-01",tz="Australia/Sydney")
dateEnd <- ymd("2023-05-05",tz="Australia/Sydney")
numDays <- as.integer(dateEnd - dateStart + 1)

occurrenceDays <- filter(df.alltags,dateAus >= dateStart, dateAus <= dateEnd) %>% 
  count(dateAus,motusTagID) %>%
  count(motusTagID) %>%
  rename("days"="n") %>% 
  mutate(occurrence = days/numDays)
```

## Tidal categorisation
```{r}
# Import test data
tideTest <- read.csv("tideTest.csv")

# Convert dates to POSIXct format
tideTest$dateTimeAus <- ymd_hms(tideTest$dateTimeAus, tz = "Australia/Sydney")

tideTest$tideCategory[which.min(abs(tideTest$dateTimeAus-ymd_hms("2023-01-01 07:14:00", tz = "Australia/Sydney")))]

# This one takes ages
df.alltags <- df.alltags  %>% rowwise() %>% mutate(tideCategory = get.tideCategory(timeAus)) %>% ungroup()

```

## Occurence by day by tag by tide
```{r}
dateStart <- ymd("2023-01-01",tz="Australia/Sydney")
dateEnd <- ymd("2023-05-05",tz="Australia/Sydney")
numDays <- as.integer(dateEnd - dateStart + 1)

occurrenceDaysTide <- filter(df.alltags,dateAus >= dateStart, dateAus <= dateEnd)
occurrenceDaysTide %>% 
  count(dateAus,motusTagID,tideCategory) %>%
  count(motusTagID,tideCategory) %>%
  rename("days"="n") %>% 
  mutate(occurrence = days/numDays)






