---
title: "MOTUS telemetry array to access habitat selection from migratory shorebirds within and across Hunter and Port Stephen estuaries, NSW"
date: last-modified
date-format: "MMMM D, YYYY"
css: styles.css
format:
  html:
    code-line-numbers: true
resources:
  - figures/*
project:
  type: website
  output-dir: docs
---


<button
  class="backBtn"
  data-target="contents4" 
  onclick="document.getElementById(this.getAttribute('data-target')).scrollIntoView({behavior: 'smooth'});"
  style="position: fixed; bottom: 20px; right: 20px; z-index: 9999; padding: 10px 15px; background-color: #007bff; border: none; border-radius: 5px; color: white; cursor: pointer; box-shadow: 0 2px 4px rgba(0,0,0,0.3); display: none;">
  Back to Contents
</button>


## 5. Shorebirds behaviour

### [Contents]{#contents5 style="color: grey; font-size: smaller;"}

-   [Packages](#packages5)
-   [Load the data](#load-the-data5)
-   [Adding crucial variables](#adding5)
-   [Plotting results](#plotting5)

---

### Packages {#packages5}

```{r 5 packages, message = FALSE, warning = FALSE, eval = TRUE}

library(dplyr)
library(here)
library(forcats) 
library(lubridate)
library(purrr) 
library(lubridate)
library(ggplot2)
library(stringr)

```

### Load the data {#load-the-data5}

Either load your own data generated across the previous chapters (.rds format), or download the data up to the-last-document-update date provided here below (contact maxime.marini\@uon.edu.au for password access):

Receiver information: <a href="#" onclick="return checkPassword('recv')">recv-info.rds</a><br>
Bird data: <a href="#" onclick="return checkPassword('bird')">data.rds</a>

```{=html}
<script>
  function checkPassword(type) {
    var correctPasswords = {
      recv: "provence13",
      bird: "provence13"
    };
    var urlLinks = {
      recv: "https://uoneduau-my.sharepoint.com/:u:/g/personal/c3541851_uon_edu_au/EemdZD3VJUdHvkVoKbqUhsABCJekUx7zYJ-_5snIw0BI4A?e=RRnDGy",
      bird: "https://uoneduau-my.sharepoint.com/:u:/g/personal/c3541851_uon_edu_au/EU1v7RPMqwZGi-fapl0G620BYQVUbAQcr1tHyDqgsEPBKg?e=pfZgDd"
    };
    var userPassword = prompt("Please enter the password to access the link:");
    if (userPassword === correctPasswords[type]) {
      window.open(urlLinks[type], '_blank');
    } else {
      alert("Incorrect password. Access denied.");
    }
    return false; // prevent default link behavior
  }
</script>
```

To properly load the data into your environment, make sure to set the working directory and the path as yours.

```{r 5 load data 1, eval = FALSE}
# Set WD
setwd(dirname(rstudioapi::getSourceEditorContext()$path)) 
# Birds
data_all <- readRDS(here::here("data", "recv.info.rds"))
# Receivers info
recv <- readRDS(here::here("data", "data.rds")) 
```

```{r 5 load data 2, echo = FALSE, eval = TRUE}

# Birds
data_all <- readRDS(
  tail(sort(list.files(
  here::here("1_data", "alltags", "motus.rds"),
  pattern = "-data\\.rds$", full.names = TRUE
  )), 1)) # pick up the most recent .rds file

# Receivers info
recv <- readRDS(
  tail(sort(list.files(
    here::here("1_data", "alltags", "motus.rds"),
    pattern = "-recv-info\\.rds$", full.names = TRUE
  )), 1)) 

# Selecting upon our needs
data <- data_all %>%
  select(Band.ID, recvDeployName, recv, speciesEN, sig, sigsd, sigPositive, tideCategory, timeAus)

```

### Adding crucial variables {#adding5}

```{r 5 compute variables, echo = TRUE, eval = TRUE}

# Mean signal strengh per individual, site and time categories (HD, HN, LD, LN)
data <- data %>%
  group_by(Band.ID, tideCategory, recvDeployName) %>%
  mutate(sig_indiv_tideCat = mean(sigPositive)) %>%
  ungroup()  %>%
  mutate(speciesINIT = str_split(speciesEN, " ") %>%            
           map_chr(~ str_c(str_sub(.x, 1, 1), collapse = "")) %>%    
           toupper(),
         tideCat = str_split(tideCategory, "_") %>%          
           map_chr(~ str_c(str_sub(.x, 1, 1), collapse = "")) %>%    
           toupper())

```

Here, we take the raw values of signal strength `sig` to calculate the mean for each individual and per tide category `tideCategory`. So we can group the means per species and per site to observ how the signal strength varies. To that aim, we need to create a new variable: the average signal strength per individuals `sig_indiv_tideCat`.

### Plotting results {#plotting5}

Below is displayed the variance of the signal strength among species and Motus stations.Hypotesising that foraging activity (high variance values) would more occur in Low tide context, or/and at night.

```{r 5 plotting results , echo = FALSE, eval = TRUE}

# Color codes
species_colors <- c(
  "Bar-tailed Godwit"      = "#1b9e77",  
  "Far Eastern Curlew"     = "#d95f02",  
  "Masked Lapwing"         = "#7570b3", 
  "Pacific Golden-Plover"  = "#e7298a", 
  "Pied Stilt"             = "#66a61e", 
  "Red-necked Avocet"      = "#e6ab02"   
)

speciesINIT_colors <- c(
  "BG"  = "#1b9e77", 
  "FEC" = "#d95f02",  
  "ML"  = "#7570b3",  
  "PG"  = "#e7298a",  
  "PS"  = "#66a61e", 
  "RA"  = "#e6ab02"   
)

```

```{r 5 plotting results2 , echo = TRUE, eval = FALSE}
# One plot per species for better understanding
data_split <- split(data, data$speciesEN) 
plots_list <- lapply(names(data_split), function(species) {
  df_sub <- data_split[[species]]
  ggplot(df_sub, aes(x = tideCat, y = sig_indiv_tideCat, fill = speciesEN)) +
    geom_boxplot() +
    scale_fill_manual(values = species_colors, guide = "none") +
    facet_wrap(~ recvDeployName) +
    theme_bw() +
    labs(
      title = paste0("Signal Strength variance for ", species, " over MOTUS stations"),
      x = "Tide Category",
      y = "Signal Strength variance (RSSI)",
      fill = "Species"
    )
})
```

::: panel-tabset
```{r 5 plotting results2BIS, echo = FALSE, eval = TRUE, results = 'asis'}
# One plot per species for better understanding
plots_list <- data %>%
  split(.$speciesEN) %>%
  map(~ ggplot(.x, aes(x = tideCat, y = sig_indiv_tideCat, fill = speciesEN)) +
        geom_boxplot() +
        facet_wrap(~ recvDeployName) +
        scale_fill_manual(values = species_colors, guide = "none") +
        theme_bw() +
        labs(
          title = paste0("Signal Strength variance for ", unique(.x$speciesEN), " over MOTUS stations"),
          x = "Tide Category",
          y = "Signal Strength variance (RSSI)",
          fill = "Species"
        ))
```

```{r 5 Plotting results2BIS, echo = FALSE, eval = TRUE, results = 'asis'}

# Use knitr::asis_output to ensure output is captured
for (species_name in names(plots_list)) {
  cat("\n\n### ", species_name, "\n\n")
  print(plots_list[[species_name]])
  flush.console() #force immediate output
}
```
:::


<button
  class="backBtn"
  data-target="contents5" 
  onclick="document.getElementById(this.getAttribute('data-target')).scrollIntoView({behavior: 'smooth'});"
  style="position: fixed; bottom: 20px; right: 20px; z-index: 9999; padding: 10px 15px; background-color: #007bff; border: none; border-radius: 5px; color: white; cursor: pointer; box-shadow: 0 2px 4px rgba(0,0,0,0.3); display: none;">
  Back to Contents
</button>

::::::

<script>
document.addEventListener("DOMContentLoaded", function() {
  const buttons = Array.from(document.querySelectorAll(".backBtn"));
  window.addEventListener("scroll", function() {
    let scrollPos = window.pageYOffset || 
                    document.documentElement.scrollTop || 
                    document.body.scrollTop || 0;
    let buttonToShow = null;
    buttons.forEach(function(btn) {
      let targetId = btn.getAttribute("data-target");
      if (!targetId) return;
      let targetEl = document.getElementById(targetId);
      if (!targetEl) return;
      if (scrollPos > targetEl.offsetTop + 100) {
        if (!buttonToShow || targetEl.offsetTop > document.getElementById(buttonToShow.getAttribute("data-target")).offsetTop) {
          buttonToShow = btn;
        }
      }
    });
    buttons.forEach(function(btn) {
      if (btn === buttonToShow) {
        btn.style.display = "block";
      } else {
        btn.style.display = "none";
      }
    });
  });
});
</script>



***Availability***

*The analysis are ran through R version 4.5.1 (2025-06-13 ucrt) -- "Great Square Root" and the platform: x86_64-w64-mingw32/x64.*

```{r check, echo = FALSE, eval = FALSE}

############ /!\ always check /!\ #############

# spreadsheet up to date
# downloadable data into the share-point up to date + the link to access matching the up to date data file
# once the filtering done, check every argument being ID when referencing to individuals


```
